---
title: "Olayinka_Arimoro_R_assignment"
author: "Olayinka Arimoro"
date: "December 9, 2019"
output: html_document

This assignment was completed on `r date()`. I successfully made the publication, here is the link http://rpubs.com/Predicare/560226 to the publication

```{r Setting working directory}
# Setting the working directory
setwd("C:/Users/hp 2018/Desktop/Data_Science/iAspire_Fellow/Module_One/Essential_Statistics_and_R/R/student_files_r")
```
### a
```{r}
datamininggrades=c(90,87,69,89,58,93,99,98,76,88)
sort(x=datamininggrades)
```
### b
```{r}
sort(x=datamininggrades, decreasing = TRUE)
```
### c
```{r}
lettergrade <- c("A-", "B+", "F", "B+", "F", "A", "A", "A", "C+", "B+")
sort(x=lettergrade)
```
### d
```{r}
lettergradefactor <- factor(lettergrade)
lettergradefactor
# Sorting lettergradefactor
sort(x=lettergradefactor)
```
### e (Sorting an ordinal variable)
```{r}
#create the variable
agreement <- c("Strongly Disagree", "Disagree", "Somewhat Disagree","Neither agree or disagree", "Somewhat agree", "Agree", "Strongly Agree")
#view the variable
agreement

# convert to factor
agreementfactor <- factor(agreement)
# View the factor created
agreementfactor

# Sort agreement factor
sort(x=agreementfactor)
# Observation: It sorted using the alphabets, which is not really useful

# Sorting by factor levels
# Creating a factor of "agreement" and defining its levels
agreementfactor <- factor(agreement, levels = c("Strongly Disagree", "Disagree", "Somewhat Disagree","Neither agree or disagree", "Somewhat agree", "Agree", "Strongly Agree"))
# sort
sort(x=agreementfactor)
```
### f (Creating a dataframe)
```{r}
finalgrades <- data.frame(datamininggrades, lettergrade)
finalgrades
```
### g (Sorting a dataframe)
```{r}
# In ascending order
finalgrades[with(finalgrades, order(datamininggrades)), ]
# In descending order
finalgrades[with(finalgrades, order(-datamininggrades)), ]
```
### h (Reshaping a dataframe)
```{r}
# Importing the dataset
internet <- read.csv("world_internet_usage.csv", header=T, sep=",")
# Viewing the dataset
View(internet)

# Change the column names
names(internet) <- c("country", "2000", "2001", "2002", "2003", "2004", "2005", "2006", "2007", "2008", "2009", "2010", "2011", "2012")

# Checking the structure of the data "internet"
str(internet)
```
### i (Creating a histogram for a year)
```{r}
# single histogram for a single year
hist(internet$"2000", breaks=8, freq=TRUE, 
    main="Histogram of Internet Usage in 2000", 
    xlab = "Internet Usage")
```
### j (Plotting a matrix of histograms)
```{r}
require(lattice)
h1 =histogram(internet$"2000", breaks=8, xlab="Internet Usage", main="2000")
h2 =histogram(internet$"2001", breaks=8, xlab="Internet Usage", main="2001")
h3 =histogram(internet$"2002", breaks=8, xlab="Internet Usage", main="2002")
h4 =histogram(internet$"2003", breaks=8, xlab="Internet Usage", main="2003")
h5 =histogram(internet$"2004", breaks=8, xlab="Internet Usage", main="2004")
h6 =histogram(internet$"2005", breaks=8, xlab="Internet Usage", main="2005")
h7 =histogram(internet$"2006", breaks=8, xlab="Internet Usage", main="2006")
h8 =histogram(internet$"2007", breaks=8, xlab="Internet Usage", main="2007")
h9 =histogram(internet$"2008", breaks=8, xlab="Internet Usage", main="2008")
h10 =histogram(internet$"2009", breaks=8, xlab="Internet Usage", main="2009")
h11 =histogram(internet$"2010", breaks=8, xlab="Internet Usage", main="2010")
h12 =histogram(internet$"2011", breaks=8, xlab="Internet Usage", main="2011")
h13 =histogram(internet$"2012", breaks=8, xlab="Internet Usage", main="2012")

print(h1, split=c(1,1,4,4), more=T);
print(h2, split=c(2,1,4,4), more=T);
print(h3, split=c(3,1,4,4), more=T);
print(h4, split=c(4,1,4,4), more=T);
print(h5, split=c(1,2,4,4), more=T);
print(h6, split=c(2,2,4,4), more=T);
print(h7, split=c(3,2,4,4), more=T);
print(h8, split=c(4,2,4,4), more=T);
print(h9, split=c(1,3,4,4), more=T);
print(h10, split=c(2,3,4,4), more=T);
print(h11, split=c(3,3,4,4), more=T);
print(h12, split=c(4,3,4,4), more=T);
print(h13, split=c(1,4,4,4), more=F);
```
### k (transform internet data frame from wide to long format)
```{r}
library(reshape2)
iw_melt = melt(internet,id.vars="country", variable.name="year", value.name="usage")
View(iw_melt)
```
### l (Plotting the data )
```{r}
plot(iw_melt$year,iw_melt$usage, main="Box plot of Year and usage of internet")
```
### m (Using the ggplot function to plot a line plot for each country)
```{r}
library(ggplot2)
ggplot(iw_melt,aes(x=year,y=usage,colour=country,group=country)) + geom_line(lwd=1)
```
### n (while loops)
```{r}
counter = 0
while (counter <9) {
  print(counter)
  counter = counter + 1
}
```
### o (for loops)
```{r}
for (value in 1:10)
{
  print(value)
}
```
### p 
```{r}
for (i in internet$country)
{
	caps = toupper(i)
print (caps)
}
```
### q
```{r}
weather = "rain"
if (weather == "rain") {

#bring umbrella
print("Don't forget your umbrella")
}
```
### r
```{r}
meaninternet = mean(internet$"2012")
if (internet$"2012"[1] <= meaninternet) {
  print (as.character(internet$country[1]))
}
```
### s
```{r}
meaninternet = mean(internet$"2012")
if (internet$"2012"[1] >= meaninternet) {
  print (as.character(internet$country[1]))
} else {
  cat (as.character(internet$country[1]), "has a lower internet usage than the mean.")       
}
```
### t
```{r}
meaninternet = round(mean(internet$"2012"),2)
x = 0
for (i in internet$"2012"){
  x =  x + 1
  #print(x)
  if (internet$"2012"[x] > meaninternet){
    country = as.character(internet$country[x])
    cat(country, "is greater than the mean of", meaninternet, 
        "with a value of", internet$"2012"[x],"\n")
  }
}
```
### u
```{r}
s = "10-12-1977"
s1 = strsplit(s, split='-', fixed=TRUE)
s1

# Return the third element
s = "10-12-1977"
s1 = unlist(strsplit(s, split='-', fixed=TRUE))[3]
s1


npv <-function (Pt,r,t){
  pv=Pt/(1+r)^t
  return (pv)
}
# Printing the result
Pt<-50
r<-.05
t<-1
npv(Pt,r,t)

```
### v
```{r}
# Writing the functions
descriptives=function(x){
  c(mean     = mean(x),
    sd       = sd(x),
    median   = median(x),
    min      = min(x),
    max      = max(x),
    range    = max(x)-min(x),
    quantile = quantile(x),
    IQR      = IQR(x),
    var      = var(x)              
  )
}

#execute function below
sapply(attitude, descriptives)

```
### w
```{r}
# import march madness data
march <- read.csv('march_madness.csv', header=T, sep=',')
# change columns to lowercase
names(march) = tolower(names(march))
library(stringr)
wins=unlist(str_split(march$record, '-'))
wins
march_wins = data.frame(march,wins)
View(march_wins)

# Now, write a function for exercise w. 

library(stringr)
recordscore = function(x)
{
  c(unlist(str_split(x, '-'))[1])
  
}  

march_wins =sapply(march$record, recordscore) 
march_wins = as.numeric(march_wins)  # Converting to a numeric class
class(march_wins)  # Checking to see that it is a numeric class
max(march_wins)   # Checking the maximum value
march_new = data.frame(march$school, march_wins)

View(march_new)
names(march_new) =c("school","wins")
View(march_new)

```
### x
```{r}
# Import sidewalk dataset
sidewalk <- read.csv("Sidewalk_Cafes.csv", header=T, sep=",")

#Check class
class(sidewalk)

#mean of square footage
mean(sidewalk$Lic.Area.Sq.Ft, na.rm='TRUE')

#boxplot of the NYC slidewalk cafe size
boxplot(sidewalk$Lic.Area.Sq.Ft, main="NYC Sidewalk Cafe Size")

```
### y
```{r}
enclosed <- subset.data.frame(sidewalk, sidewalk$Sidewalk.Cafe.Type == 'Enclosed')

unenclosed <- subset.data.frame(sidewalk, sidewalk$Sidewalk.Cafe.Type == 'Unenclosed')
```
The data was then plotted with this code:
```{r}
boxplot(enclosed$Lic.Area.Sq.Ft, unenclosed$Lic.Area.Sq.Ft, names=c("Enclosed","Unenclosed"), main="NYC Sidewalk Cafe Size", horizontal = TRUE, xlab = "Square Footage")
```
### Summary for the boxplot above
```{r}
# The median size of the reported 139 enclosed cafes is 325.5 square feet. The largest enclosed cafe is 1044 square feet and the smallest is 1 square feet.

# The median size of the reported 834 enclosed cafes is 202 square feet. The largest enclosed cafe is 1300 square feet and the smallest is 30 square feet.
```
### z (Using the Shiny App)
```{r}
require(shiny)
runExample("01_hello")

library(shiny)

# Define server logic required to draw a histogram
shinyServer(function(input, output) {
  
  # Expression that generates a histogram. The expression is
  # wrapped in a call to renderPlot to indicate that:
  #
  #  1) It is "reactive" and therefore should re-execute automatically
  #     when inputs change
  #  2) Its output type is a plot
  
  output$distPlot <- renderPlot({
    sidewalk <- read.csv("Sidewalk_Cafes.csv")
    x   <-  na.omit(sidewalk$Lic.Area.Sq.Ft)
    bins <- seq(min(x), max(x), length.out = input$bins + 1)
    
    # draw the histogram with the specified number of bins
    hist(x, breaks = bins, col = 'darkgray', border = 'white', 
    main="Histogram of Square Footage of NYC Sidewalk Cafes", 
    xlab="Cafe Area Square Footage", ylab="Frequency")
  })
})

# And now ui.R
library(shiny)

# Define UI for application that draws a histogram
shinyUI(fluidPage(
  
  # Application title
  titlePanel("NYC Sidewalk Cafe Data"),
  
  # Sidebar with a slider input for the number of bins
  sidebarLayout(
    sidebarPanel(
      sliderInput("bins",
                  "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)
    ),
    
    # Show a plot of the generated distribution
    mainPanel(
      plotOutput("distPlot")
    )
  )
))

```








